Species richness and food-web structure jointly drive community biomass and its temporal stability in fish communities

### 2021-02-17 manuscript/bef_stability/manuscript.Rmd
Advance review

---
title: Species richness and network structure jointly drive total biomass and its temporal stability in fish communities
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault, Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    toc: false 
    fig_caption: true 
    keep_tex: true
fontsize: 12pt
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

@@

- Running title: Foodweb structure, biomass and stability

@@

Biodiversity-ecosystem functioning (BEF) and foodweb complexity-stability
relationships are central to ecology. However, they remain largely untested for 
foodweb in natural context. Here we study the links among
environmental conditions, richness, foodweb structure, annual biomass and its
temporal stability using a standardized monitoring of 99 stream fish communities
spanning from 1995 to 2018. We first reveal that both richness and average
trophic level are positively related to annual biomass, with effects of similar
strength. Second, we show that network structure affects community stability,
with a positive effect of mean trophic level, and contrary to general
expectation, we find that richness destabilizes fish communities. Finally, we
find that environmental conditions affect both biomass and its stability via
effects on richness and network structure. Strikingly, effects on stability are
mediated by changes in species stability rather than synchrony, which contrasts
with results from single-trophic communities and might be characteristic of
multi-trophic communities.

@@

Current biodiversity losses and further global change have the potential to
affect ecosystems in complex ways [@loreau_biodiversity_2001;
@tilman_biodiversity_2014]. Biodiversity has indeed a tremendous effect on
ecosystem functions and their stability in response to perturbations
[@hector_plant_1999; @loreau_biodiversity_2001; @hooper_effects_2005;
@dunne_network_2006; @duffy_why_2009], because it largely determines both the
amount and the distribution of matter and energy fluxes within and among
ecosystems [@loreau_meta-ecosystems:_2003]. Understanding the relationship
between biodiversity and ecosystem properties is thus crucial to better
anticipate the impacts of global change.

@@

Biodiversity affects biomass production [@tilman_productivity_1996;
@cardinale_species_2002; @grace_integrative_2016], distribution
[@thebault_relationship_2006] and stability [@thebault_trophic_2005;
@thebault_relationship_2006; @tilman_biodiversity_2006]. Indeed species richness
generally increases primary productivity and total biomass in experimental plant
communities [@tilman_productivity_1996; @grace_integrative_2016;
@hector_plant_1999; @tilman_biodiversity_2014; @grace_integrative_2016], stream
feeders, [@cardinale_species_2002] and aquatic bacterial consumers
[@pennekamp_biodiversity_2018]. In parallel, such positive relationship between
species richness and temporal stability of biomass has been reported in
experimental grasslands and aquatic bacterial systems
[@tilman_biodiversity_2006; @pennekamp_biodiversity_2018] as well as in natural
settings, such as stream fish communities [@franssen_annual_2011], bird and
butterfly communities [@olivier_urbanization_2020]. Despite the diversity of
ecosystems and taxa studied, the relationship between diversity and ecosystem
functioning remains mostly focused on
single trophic levels, ignoring the effects of foodweb structure. This gap is
even more striking for the relationship between diversity and temporal stability
, in spite of long-standing interest for the links
between foodweb complexity and stability in theoretical ecology [e.g.
@may_will_1972; @dunne_network_2006; @neutel_reconciling_2007; @duffy_why_2009;
@stouffer_compartmentalization_2011].

Overall, the evidence from theoretical models suggests that both species
richness and foodweb structure should modulate ecosystem properties such as
total biomass and its stability. Complementarity in the use of
resources, a main mechanism previously highlighted in single-trophic studies, is
important to understand the effects of species richness in foodwebs [@thebault_food-web_2003;
@poisot_trophic_2013; @schneider_animal_2016], but in that case it also depends
on foodweb connectance [@thebault_food-web_2003; @poisot_trophic_2013].

As classically found in studies considering a single trophic level,
[@tilman_biodiversity_2006; @loreau_species_2008; @olivier_urbanization_2020],
asynchrony in species temporal variations is also explaining a significant part
of the stabilizing effects of species richness on total biomass in foodwebs
[@thebault_trophic_2005]. Indeed, a low synchrony in species fluctuations,
i.e. asynchrony, is synonymous of compensatory dynamics, where the decrease in
abundance of a species is compensated by the increase of an other. However,
higher connectance can reverse the positive effects of species
richness on temporal stability of total biomass, by leading to a negative
relationship between species richness and species temporal variability
[@thebault_trophic_2005]. In addition, the presence of species at higher trophic
levels increases total primary production [@wang_biodiversity_2018]. The
addition of predators has also cascading effects on both species variability and
synchrony at lower trophic levels [@teng_dynamics_2004;
@shanafelt_stability_2018]. Overall, these theoretical studies show that the
effects of species richness and network structure, such as connectance and
average trophic level, are potentially conflicting or synergistic in such a way
that they need to be addressed jointly.

The gap in studies that consider jointly effects of diversity and foodweb
structure on ecosystem functions can be explained by the challenges associated
with manipulating both species richness and foodweb complexity in an
experimental context [e.g. @sander_ecological_2017] and by the fact that describing
trophic interactions between species is highly labour intensive. These
challenges are even more acute for the study of the stability of ecosystem
properties, which additionally requires time-series data and longer-term
analyses. In this context, recent developments to predict species trophic
interactions [@gravel_inferring_2013], combined with observational long-term surveys
of ecological communities, can offer an opportunity to investigate the links
between diversity, foodweb structure and the stability of total community
biomass in natural settings.

Such approach based on data from natural communities can, in addition, give
insights on the effects of environmental variables on ecosystem properties and
their stability. Environment affects organism performance directly, and thereby,
species abundances, total biomass [@grace_integrative_2016; @woods_testing_2020]
and its temporal stability [@oliver_heterogeneous_2010; @franssen_annual_2011;
@hansen_climate_2013; @bluthgen_land_2016; @de_boeck_patterns_2017;
@friedpetersen_drivers_2020]. For example, larger ecosystem might increase the
number of available resources and thus total biomass [@post_ecosystem_2000;
@doi_resource_2009] while temperature affects ecosystem respiration and biomass
distribution [@yvondurocher_warming_2011; @schwarz_warming_2017;
@tabi_warming_2019]. Environment also directly affect species richness and
foodweb structure. For instance, larger ecosystems host foodwebs with higher
diversity and trophic levels [@mchugh_dual_2010] while temperature can affect
foodweb diversity and structure in contrasted ways depending on the ecosystem
studied [@ogorman_unexpected_2017; @gauzens_biodiversity_2020]. In natural
communities, the relationships among species richness, total biomass and its
temporal stability might also arise from effects of the environment on both
diversity and ecosystem functions [@loreau_biodiversity_2001; @he_global_2013;
@grace_integrative_2016].

@@

Because their trophic interactions are well known, stream fish communities
constitute a good model to study the effects of network structure on total
biomass and its temporal stability. A substantial body of literature
supports that trophic interactions can be inferred through body size and diet
informations [@jennings_weak_2001; @gravel_inferring_2013], especially in stream
communities including fishes [@brose_predator_2019].

Our aim is to document the relationships among species richness, foodweb
structure, total biomass and its temporal stability in natural communities. To
do so, we used a standardized sampling of fish
communities in French streams [@edeline_ecological_2013]. We selected 99 sites
that have been sampled for more than 10 years. We inferred 1496 interaction
networks, based on body size and diet. We question whether species richness,
connectance and average trophic level are related to total biomass and its
temporal stability. We further question how ecosystem size and temperature are
related to total biomass and its temporal stability.

@@

```{r, cache = FALSE}
myload(qt_station, dir = mypath("data"))
matmet_cap <- paste0( 
  "(A) Observation sites (black dots) were located across France and monitored
  on the period 1995-2018. The red dots show the location of (C) and (D) sites.
  ", " Inner borders draw the limits of the hydrographic basins. ", "(B) The
  metaweb of the study. Each trophic species (N = 412) is represented by a node
  and each
  color represents a species (S = 45). Resource nodes (7) are drawn in grey. ",
  "The species names corresponding to the three digit codes are found in Table
  S1." ,
  "Yearly temporal series of networks of two stations having respectively a low
  (C) and high (D) temporal stability of biomass. C and D stations have
  respectively biomass stability the closest to the first and third quartile of
  the distribution of the temporal stability of biomass in the dataset,
  respectively ", round(qt_station$bm_std_stab[1],2), " and ",
  round(qt_station$bm_std_stab[2],
    2), ". Point size is proportional to the biomass of the trophic species.")

@@

Fish communities were monitored across stream sections in France on the period
1995-2018 by the French Office of Water and Aquatic Ecosystems (ONEMA) using
electrofishing. Two different standardized protocols were used for small and
large streams (detailed in appendix B, part 1.1). In the case where a stream was
sampled alternatively with one of the two sampling protocols, we kept only the
most frequently used. The sampling season spanned from late spring to autumn.
There was low heterogeneity of the sampling month within stations, with the
median sampled month being mid-August while the median of the standard deviation
of the sampled month within station was $0.7$.

We selected the stations that had been sampled ten years or more. As the concept
of community variability referred to variability around an equilibrium, which is
rarely observed or difficult to characterize in natural communities, only
stations with no temporal trends in total biomass were selected (detailed in
appendix B, part 2). However, the results were robust to the relaxation of this
hypothesis (Fig. S7). These selection steps resulted in 99 stations distributed
over seven hydrographic basins (Fig. \@ref(fig:matmet)A).

## Total biomass and its temporal stability

@@

The body size of all fish individuals was not recorded when they were too
numerous (Fig. S1, appendix B, part 1.1). The body size of the unmeasured
individuals was inferred under the assumption that species body size followed a
normal distribution (Fig. S1, appendix B, part 1.2). The body mass of each
individual was then estimated following the results of two meta-analyses on
fishes [@lleonart_removing_2000; @froese_cube_2006], using an allometric law
$B_i = 0.01 \times l_i^{3.03}$, with $B$ the body mass in grams, $l$ the body
size in millimeters and $i$ the individual. Species biomass and total biomass at
each sampling event were then calculated by summing individual body masses. To
account for the variation in sampling effort among sites, species biomass and
total biomass were corrected by the surface sampled (range `r range_surface[1]` -
`r range_surface[2]` square metres), and expressed in grams per square meter
($gm^{-2}$). In the analysis, the total biomass of a site has been defined as
the median of the total biomass across sampling years.

@@

We described the mechanisms driving the temporal stability of biomass,
decomposing the $CV$ of total biomass ($CV_{com}$) into a synchrony ($\phi$) and
a weighted average species $CV$ ($\overline{CV_{sp}}$) components according to
@thibaut_understanding_2013:

@@

with $CV_{com, i}$ being the CV of the biomass in the
station $i$, $\overline{CV_{sp, i}}$ the average CV of
species in the station $i$ weighted by the relative biomass of the species, $B_{k,i}$.
Synchrony of the station $i$, $\phi_i$, is the ratio between the variance of
total biomass over time ($\sigma^2_{x_{T,i}}$) and the sum of the variance of
population biomass over time ($\sigma_{x_k}$), $i$ being the station $i$ and $k$
the species $k$ [@loreau_species_2008].

@@

## Species richness and foodweb structure

@@

We determined the network structure in two steps: (1) build a metaweb describing the
trophic interactions between all trophic species and the resources (Fig.
\@ref(fig:matmet)B), then (2) determine network of a given community by
extracting the metaweb according to the trophic species present at a given
site and sampling event (Fig. \@ref(fig:matmet)C and D).

The trophic interaction network was inferred from body size and ontogenic diet
shift following previous studies [@edeline_ecological_2013;
@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019]. The diet
of stream fishes, as for many organisms, shifts throughout an individual's life,
and thereby with body size. For example, alvins feed on plankton while adults
may feed on algae and fishes, with bigger individuals eating bigger fishes. Each
fish species was hence divided in nine body size classes, each corresponding to
a trophic species (Fig. S2). Additionally to the trophic species, we added
seven resource nodes present in the ontogenic food diet database: detritus,
biofilm, phytoplankton, zooplankton, macrophages, phytobenthos and zoobenthos
(grey nodes in Fig. \@ref(fig:matmet)B, C & D). The trophic species and the
resources constituted the 412 nodes of the metaweb, i.e. 45 fish species divided
into nine body size classes plus seven resource nodes. The inference of trophic
interactions was shown to be more accurate when considering body size classes,
i.e. trophic species, than species [@jennings_weak_2001].

The feeding interactions of a fish node, i.e. a trophic species, were determined
by both its species identity and its body size. The ontogenic fish diet
database was filled according to fishbase and literature (appendix A, Table S2). Each
fish species had two or three life stages which are delimited by a body size
range (Fig. S2). The life stage of a fish trophic species was determined with
the center of its body size range, hereafter midpoint (Fig. S2).

For a piscivorous trophic species, the fish-fish trophic links were defined in
two steps and based on the body size ratio between predators and prey. First of
all, the predation window of a piscivorous trophic species was defined as 3% to
45% of its midpoint [@mittelbach_ontogeny_1998; @claessen_impact_2002;
@hart_handbook_2008]. Then, a trophic link was set between the piscivorous
trophic species and every trophic species whose midpoint was included in the
predation window (Fig. S2). The fish-resource trophic links were set between
the trophic species and the resources nodes according the food diet of their
size class.  Lastly, the trophic links among resource nodes were set according
to the literature [@allan_stream_2007; @hart_handbook_2008].

@@

For each trophic network, corresponding to a particular site and year, we
computed the connectance and mean trophic level. Connectance describes the
level of generalism in the network and is calculated as the actual number of
interactions divided by the number of interactions if all species interact
together ($C = L / N^2$, L and N being resp. the number of links and of nodes).
The trophic level of the nodes was computed as one added to the average trophic
level of its food item. The mean trophic level $\overline{T}$ was 
weighted by the biomass of each trophic species (excluding resource nodes for
which we had no biomass information), so $\overline{T} = \sum_i^N{B_iT_i \times
(\sum_i^N{B_i})^{-1}}$, with $T_i$ and $B_i$ being respectively the trophic
level and the biomass of the node $i$. For each site, the median of connectance
and average trophic level over time was used in subsequent analysis.

pca_rotated <- compute_rotated_pca(
  .data = habitat_press[, mask],
  naxis = 2
)
habitat_pressure <- 
  cbind(habitat_press[, c("station", "temperature_med", "alt")], pca_rotated$rotated$score) %>%
  as_tibble
ggplot(habitat_pressure, aes(x = RC2, y = temperature_med)) +
  geom_point()

ggplot(
  pivot_longer(habitat_pressure, cols = c("RC1", "RC2"),
    names_to = "variable", values_to = "value"
  )
  , aes(x = value)) +
geom_histogram() +
facet_grid(cols = vars(variable))

mysave(habitat_press, habitat_pressure, dir = mypath("data"), overwrite = TRUE)

@@

We characterised the sites by their altitude, slope, distance from
source, stream width and depth and its strahler order, i.e. its position in the
stream network. We obtained data on water temperature and flow from the naiades
database
[(www.naiades.eaufrance.fr)](http://www.naiades.eaufrance.fr/). As data from naiades are
not collected on the same site as our fish sampling sites, we performed an
interpolation of the water temperature and flow based on the spatial stream
network [@isaak_applications_2014; @hoef_ssn:_2014]. The interpolation procedure
is detailed in appendix B (part 1.4).

We had one environmental variable value for each sampling event, except
for water temperature for which the record began in 2006. Then, we computed the
average and CV values of water temperature, flow, Biological Oxygen Demand (BOD), river width and
depth. As habitat and environmental variables are often collinear, a principal
component analysis (PCA) was performed on the 14 scaled variables. We kept in
the analysis the two first principal components based on the elbow method, on
the eigenvalues distribution (Fig. S4). The two first principal components
explained `r tot_var_expl_pca`% of the variance (Fig. S4). A varimax axis rotation was
performed to maximize the representation of the environmental variables by the principal
components [@kaiser_varimax_1958; @revelle_psych_2019]. The first axis
represented a gradient from small to big streams, characterized by their width
and depth, but also a gradient from stream close to far from the source. We
reversed the values of the second axis, such as it represented a gradient of
increasing average water temperature, BOD and decreasing altitude and slope.
Further description of the variables are provided in appendix B (part 1.4).

@@

Structural Equation Modeling enables to assess multivariate relationships
between variables. It is a promising way to go beyond bivariate relationships
and to investigate complex relationships, direct and indirect, between variables
[@grace_structural_2008; @grace_causal_2014; @lefcheck_piecewisesem_2016;
@grace_integrative_2016]. We defined a structural equation model explaining the
average annual total biomass as a function of the two synthetic environmental
variables, species richness and foodweb structure [@maureaud_biodiversityecosystem_2019;
@woods_testing_2020]. We assumed that both species richness and network
structure directly affect total biomass (Fig. \@ref(fig:sem-plot)A)
and that species richness affected foodweb structure [@winemiller_must_1989;
@dunne_network_2006; @thebault_stability_2010].

We used the same model structure for temporal stability of biomass (Fig.
\@ref(fig:sem-plot)C). Instead of the annual total biomass, we modelled the
synchrony and $\overline{CV_{sp}}$ which in turn determined the temporal
stability of biomass.

The structural equation models were based on Gaussian linear models. All variables
except the connectance and average trophic level
were logged to fulfill model assumptions (see residual plots and data
transformation in Fig. S5 and S6, Table S5). We set the seven
hydrographic basins as a random effect on the intercept to
account for their heterogeneity. We checked for the absence of multicollinearity 
with Variance Inflation Factor (VIF, Table S6). The slopes reported
in the main text were standardized in order to compare their magnitude ($r_\delta
= \beta \frac{\sigma_x}{\sigma_y}$, $\beta$ being the unstandardized slope,
$\sigma_x$ and $\sigma_y$ being respectively the standard deviation of the
predictive variable $x$ and of the response variable $y$). The paths whose
slope had a $P$ value inferior to $0.05$ were reported in the main text, all the
path values are presented in Table S7 and S8.

>>>>

---
title: Species richness and food-web structure jointly drive community biomass and its temporal stability in fish communities
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault, Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    toc: false 
    fig_caption: true 
    keep_tex: true
fontsize: 12pt
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

@@

- Running title: food-web structure, biomass and stability

@@

Biodiversity-ecosystem functioning (BEF) and food-web complexity-stability
relationships are central to ecology. However, they remain largely untested in natural context. Here we study the links among
environmental conditions, richness, food-web structure, annual biomass and its
temporal stability using a standardized monitoring dataset of 99 stream fish communities
spanning from 1995 to 2018. We first reveal that both richness and average
trophic level are positively related to annual biomass, with effects of similar
strength. Second, we show that food-web structure affects community stability,
with a positive effect of mean trophic level, and contrary to
expectation, a destabilizing effect of species richness. Finally, we
find that environmental conditions affect both biomass and its stability via
effects on richness and network structure. Strikingly, effects on stability are
mediated by changes in population stability rather than synchrony, which contrasts
with results from single-trophic communities and might be characteristic of
multi-trophic communities.

@@

Current biodiversity losses and further global change have the potential to
affect ecosystems in complex ways [@loreau_biodiversity_2001;
@tilman_biodiversity_2014]. Biodiversity has indeed a major effect on
ecosystem functions and their stability in response to perturbations
[@hector_plant_1999; @loreau_biodiversity_2001; @hooper_effects_2005;
@dunne_network_2006; @duffy_why_2009], because it largely determines the
amount and the distribution of matter and energy fluxes within and among
ecosystems [@loreau_meta-ecosystems:_2003]. Understanding the relationship
between biodiversity and ecosystem properties is thus crucial to better
anticipate the impacts of global change.

@@

Biodiversity affects biomass production [@tilman_productivity_1996;
@cardinale_species_2002; @grace_integrative_2016], distribution
[@thebault_relationship_2006] and stability [@thebault_trophic_2005;
@thebault_relationship_2006; @tilman_biodiversity_2006]. Indeed species richness
generally increases primary productivity and total biomass in experimental plant
communities [@tilman_productivity_1996; @grace_integrative_2016;
@hector_plant_1999; @tilman_biodiversity_2014; @grace_integrative_2016], stream
feeders, [@cardinale_species_2002] and aquatic bacterial consumers
[@pennekamp_biodiversity_2018]. In parallel, such positive relationship between
species richness and temporal stability of biomass has been reported in
experimental grasslands and aquatic bacterial systems
[@tilman_biodiversity_2006; @pennekamp_biodiversity_2018] as well as in natural
settings, such as stream fish communities [@franssen_annual_2011], bird and
butterfly communities [@houlahan_negative_2017; @olivier_urbanization_2020].
Despite the diversity of ecosystems and taxa studied, the relationship between
diversity and ecosystem functioning remains mostly focused on single trophic
levels, ignoring the effects of food-web structure. This gap is even more
striking for the relationship between diversity and temporal stability , in
spite of long-standing interest for the links between food-web complexity and
stability in theoretical ecology [e.g.  @may_will_1972; @dunne_network_2006;
@neutel_reconciling_2007; @duffy_why_2009; @stouffer_compartmentalization_2011].

Overall, the evidence from theoretical models suggests that both species
richness and food-web structure should modulate ecosystem properties such as
community biomass and its stability. Complementarity in the use of
resources is important to understand the effects of species richness in food-webs
[@thebault_food-web_2003; @poisot_trophic_2013; @schneider_animal_2016], but it
also depends on food-webs connectance, i.e. its level of generalism
[@thebault_food-web_2003; @poisot_trophic_2013].

As classically found in studies considering a single trophic level,
[@tilman_biodiversity_2006; @loreau_species_2008; @olivier_urbanization_2020],
asynchrony in population temporal variations also explains a significant part
of the stabilizing effects of species richness on community biomass in food-webs
[@thebault_trophic_2005]. Indeed, a low synchrony in population fluctuations,
i.e. asynchrony, is synonymous of compensatory dynamics, where the decrease in
abundance of a population is compensated by the increase of another. However,
higher connectance can counterbalance the positive effects of species
richness on temporal stability of community biomass, by leading to a negative
relationship between species richness and population temporal variability
[@thebault_trophic_2005]. In addition, the presence of species at higher trophic
levels increases total primary production, through stronger top-down effects
associated with higher nutrient uptake [@wang_biodiversity_2018]. The addition of predators has
also cascading effects on both population variability and
synchrony at lower trophic levels [@teng_dynamics_2004;
@shanafelt_stability_2018]. Overall, these theoretical studies show that the
effects of species richness and food-web structure, such as connectance and
average trophic level, are potentially conflicting or synergistic in such a way
that they need to be addressed jointly.

The lack of studies that consider jointly effects of diversity and food-web
structure on ecosystem functions can be explained by the challenges associated
with manipulating both species richness and food-web complexity in an
experimental context [e.g. @sander_ecological_2017] and by the fact that
describing trophic interactions between species is highly labour intensive.
These challenges are even more acute for the study of the stability of ecosystem
properties, which additionally requires longer-term
perspectives. In this context, recent developments to predict species trophic
interactions [@beckerman_foraging_2006; @gravel_inferring_2013; @portalier_foraging_2019], combined with observational long-term
surveys of ecological communities, can offer the opportunity to investigate the
links between diversity, food-web structure and the stability of total community
biomass in natural settings.

Such approach based on data from natural communities can, in addition, give
insights on the effects of environmental variables on ecosystem properties and
their stability. Environment affects organism performance directly, and thereby,
population abundances, community biomass [@grace_integrative_2016; @woods_testing_2020]
and its temporal stability [@oliver_heterogeneous_2010; @franssen_annual_2011;
@hansen_climate_2013; @bluthgen_land_2016; @de_boeck_patterns_2017;
@friedpetersen_drivers_2020]. For example, larger ecosystem might increase the
number of available resources and thus community biomass [@post_ecosystem_2000;
@doi_resource_2009] while temperature affects ecosystem respiration and biomass
distribution [@yvondurocher_warming_2011; @schwarz_warming_2017;
@tabi_warming_2019]. Environment also directly affect species richness and
food-web structure, as larger ecosystems host food-webs with higher
diversity and trophic levels [@mchugh_dual_2010] whereas temperature can affect
food-web diversity and structure in contrasted ways depending on the ecosystem
studied [@ogorman_unexpected_2017; @gauzens_biodiversity_2020]. In natural
communities, the relationships among species richness, community biomass and its
temporal stability might also arise from effects of the environment on both
diversity and ecosystem functions [@loreau_biodiversity_2001; @he_global_2013;
@grace_integrative_2016].

@@

Because their trophic interactions are well known, stream fish communities
constitute a good model to study the effects of network structure on total
biomass and its temporal stability. A substantial body of literature
supports that trophic interactions can be inferred through body size and diet
informations [@jennings_weak_2001; @beckerman_foraging_2006;
@gravel_inferring_2013; @brose_predator_2019; @portalier_foraging_2019].

Our aim is to document the relationships among species richness, food-web
structure, community biomass and its temporal stability in natural communities. To
do so, we used a standardized sampling of fish
communities in French streams [@edeline_ecological_2013]. We selected 99 sites
that have been sampled for more than 10 years. We inferred 1496 interaction
networks, based on body size and diet. We question whether species richness,
food-web connectance and average trophic level are related to community biomass and its
temporal stability. We further assess how ecosystem size and temperature
mediated this relationship.

@@

pca_rotated <- compute_rotated_pca(
  .data = habitat_press[, mask],
  naxis = 2
)
habitat_pressure <- 
  cbind(habitat_press[, c("station", "temperature_med", "alt")], pca_rotated$rotated$score) %>%
  as_tibble

mysave(habitat_press, habitat_pressure, dir = mypath("data"), overwrite = TRUE)
```

```{r, cache = FALSE}
pca_caption <- paste0(
  "(B) Rotated axis of PCA on environmental variables. The five components
  explained,", tot_var_expl_pca, "% of the variance (",
  paste0(names(prop_var_expl_pca), ":", prop_var_expl_pca, "%"), ").",
  " Axis 1 was positively related to the stream size, axis 2 was related to
  the stream slope and altitude and negatively related to average water temperature and
  BOD."
  )

matmet_cap <- paste0( 
  "(A) Observation sites (black dots) were located across France and monitored
  on the period 1995-2018. The red dots show the location of (C) and (D) sites.
  ", " Inner borders draw the limits of the hydrographic basins. ", pca_caption,
  " (C) and (D) The temporal dynamic of population biomass with three networks
  of years 1997, 1999 and 2002. Each trophic species is represented
  by a node and each color represents a species. The node size are proportional to
  the biomass of the trophic species. The seven resource nodes are drawn in grey. ",
  "The species names corresponding to the three digit codes are found in Table
  S1.")

@@

Fish communities were monitored across stream sections in France on the period
1995-2018 by the French Office of Water and Aquatic Ecosystems (ONEMA) using
electrofishing. Two different standardized protocols were used for small and
large streams (detailed in appendix B, part 1.1). In the case where a stream was
sampled alternatively with one of the two sampling protocols, we only included
data derived from the most frequently used sampling method. The sampling season
spanned from late spring to autumn.  There was low heterogeneity of the sampling
month within stations, with the median sampling month being mid-August where the
median standard deviation of the sampling month was $0.7$ month, i.e. three weeks.

We selected the stations that had been sampled ten years or more. As the concept
of community variability referred to variability around an equilibrium, which is
rarely observed or difficult to characterize in natural communities, only
stations with no temporal trends in community biomass were selected (detailed in
appendix B, part 2). However, the results were robust to the relaxation of this
hypothesis (Fig. S7). These selection steps resulted in 99 stations distributed
over seven hydrographic basins (Fig. \@ref(fig:matmet)A).

## Community biomass and its temporal stability

@@

The body size of all fish individuals was not recorded when they were too
numerous (Fig. S1, appendix B, part 1.1). The body size of the unmeasured
individuals was inferred under the assumption that species body size followed a
normal distribution (Fig. S1, appendix B, part 1.2). The body mass of each
individual was then estimated following the results of two meta-analyses on
fishes [@lleonart_removing_2000; @froese_cube_2006], using an allometric law
$B_i = 0.01 \times l_i^{3.03}$, with $B$ the body mass in grams, $l$ the body
size in millimeters and $i$ the individual. Population biomass and community biomass at
each sampling event were then calculated by summing individual body masses. To
account for the variation in sampling effort among sites, population biomass and
community biomass were corrected by the surface sampled (range `r range_surface[1]` -
`r range_surface[2]` square metres), and expressed in grams per square meter
($gm^{-2}$). In the analysis, the community biomass of a site was defined as
the median of the community biomass across sampling years.

@@

We described the mechanisms driving the temporal stability of biomass,
decomposing the $CV$ of the community biomass ($CV_{com}$) into a
synchrony ($\phi$) component and a weighted average species $CV$ ($\overline{CV_{sp}}$)
component according to @thibaut_understanding_2013:

@@

with $CV_{com, i}$ being the CV of the community biomass in the
station $i$, $\overline{CV_{sp, i}}$ the average CV at the species-level in the station $i$, weighted by the relative biomass of the species, $B_{k,i}$.
Synchrony of the station $i$, $\phi_i$, is the ratio between the variance of
community biomass over time ($\sigma^2_{x_{T,i}}$) and the sum of the variance of
population biomass over time ($\sigma_{x_{k,i}}$), $i$ being the station $i$ and $k$
the species $k$ [@loreau_species_2008].

@@

## Species richness and food-web structure

@@

We determined the network structure in two steps: (1) build a metaweb describing the
trophic interactions between all fishes and the resources (Fig.
\@ref(fig:matmet)B), then (2) determine network of a given community by
extracting the metaweb according to the fishes present at a given
site and sampling event (Fig. \@ref(fig:matmet)C and D).

The trophic interaction network was inferred from body size and ontogenic diet
shift following previous studies [@edeline_ecological_2013;
@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019]. The diet
of stream fishes, as for many organisms, shifts throughout an individual's life,
and thereby with body size. For example, larvae and small juveniles feed on plankton while adults
may feed on algae and fishes, with bigger individuals eating bigger fishes. Each
fish species was hence divided in nine body size classes, each corresponding to
a trophic species (appendix B, Fig. S2). Additionally to the trophic species, we added
seven resource nodes present in the ontogenic food diet database: detritus,
biofilm, phytoplankton, zooplankton, macrophages, phytobenthos and zoobenthos
(grey nodes in Fig. \@ref(fig:matmet)B, C & D). The trophic species and the
resources constituted the 412 nodes of the metaweb, i.e. 45 fish species divided
into nine body size classes plus seven resource nodes. The inference of trophic
interactions was shown to be more accurate when considering body size classes,
i.e. trophic species, than species [@jennings_weak_2001].

The feeding interactions of a fish node, i.e. a trophic species, were determined
by both its species identity and its body size. The ontogenic fish diet
database was filled according to fishbase and literature (appendix A, Table S2). Each
fish species had two or three life stages which are delimited by its body size
range (appendix B, Fig. S2). The life stage of a fish trophic species was determined by 
the center of its body size range, hereafter midpoint (appendix B, Fig. S2).

For a piscivorous trophic species, the fish-fish trophic links were defined in
two steps and based on the body size ratio between predators and prey. First of
all, the predation window of a piscivorous trophic species was defined as 3% to
45% of its midpoint [@mittelbach_ontogeny_1998; @claessen_impact_2002;
@hart_handbook_2008]. Then, a trophic link was set between the piscivorous
trophic species and every trophic species whose midpoint in bodysize range was
included in the predation window (appendix B, Fig. S2). The fish-resource trophic links were
set between the trophic species and the resources nodes according the food items
of their size class.  Lastly, the trophic links among resource nodes were set
according to the literature [@allan_stream_2007; @hart_handbook_2008].

@@

For each trophic network, corresponding to a particular site and year, we
computed the connectance and mean trophic level. Connectance describes the
level of generalism in the network and is calculated as the actual number of
interactions divided by the number of interactions if all species interact
($C = L / N^2$, L and N being resp. the number of links and of nodes).
The trophic level of the nodes was computed as one added to the average trophic
level of its food item. The mean trophic level $\overline{T}$ was 
weighted by the biomass of each trophic species (excluding resource nodes for
which we had no biomass information), so $\overline{T} = \sum_i^N{B_iT_i \times
(\sum_i^N{B_i})^{-1}}$, with $T_i$ and $B_i$ being respectively the trophic
level and the biomass of the node $i$. For each site, the median of connectance
and average trophic level over time was used in subsequent analysis.

@@

We characterised the sites by their altitude, slope, distance from
source, stream width, depth and strahler order, i.e. its position in the
stream network hierarchy. We obtained data on water temperature, flow and Biological Oxygen
Demand (BOD) from the naiades
database
[(www.naiades.eaufrance.fr)](http://www.naiades.eaufrance.fr/). As data from naiades are
not collected on the same site as our fish sampling sites, we performed an
interpolation of the water temperature and flow based on the spatial stream
network models [@isaak_applications_2014; @hoef_ssn:_2014]. The interpolation procedure
is detailed in appendix B (part 1.4).

We had one environmental variable value for each sampling event, except
for water temperature for which the record began in 2006. Then, we computed the
average and temporal CV values of water temperature, flow, BOD, river width and
depth. As habitat and environmental variables are often collinear, a principal
component analysis (PCA) was performed on the 14 scaled variables. We kept in
the analysis the two first principal components based on the elbow method, on
the eigenvalues distribution (appendix B, Fig. S4). The two first principal components
explained `r tot_var_expl_pca`% of the variance (appendix B, Fig. S4). A varimax axis rotation was
performed to maximize the representation of the environmental variables by the principal
components [@kaiser_varimax_1958; @revelle_psych_2019]. The first axis
represented a gradient from small to big streams, characterized by their width
and depth, but also a gradient from stream close to far from the source. We
reversed the values of the second axis, such as it represented a gradient of
increasing average water temperature, average BOD and decreasing altitude and slope.
Further description of the variables are provided in appendix B (part 1.4).

@@

We analyzed the relationships among environment, species richness, food-web
structure, community biomass and its temporal stability with structural equation
models as they enable the assessment of potential indirect and direct causal
relationships among multiple variables [@grace_structural_2008;
@grace_causal_2014; @lefcheck_piecewisesem_2016; @grace_integrative_2016].  We
defined a structural equation model explaining the median annual community
biomass as a function of the two synthetic environmental variables, species
richness and food-web structure [@maureaud_biodiversityecosystem_2019;
@woods_testing_2020]. We assumed that both species richness and food-web
structure directly affect community biomass (Fig. \@ref(fig:sem-plot)A) and that
species richness affected food-web structure [@winemiller_must_1989;
@dunne_network_2006; @thebault_stability_2010].

We used the same model structure to assess the drivers of temporal stability
of community biomass (Fig.  \@ref(fig:sem-plot)C). Instead of the
annual community biomass, we modelled the synchrony and $\overline{CV_{sp}}$ which
in turn determined the temporal stability of community biomass.

The structural equation models were based on Gaussian linear mixed-effects models. All variables
except the connectance and average trophic level
were log-transformed to fulfill model assumptions (see residual plots and data
transformation in Fig. S5 and S6, Table S5, appendix B). We set the seven
hydrographic basins as a random effect on the intercept to
account for their heterogeneity. We checked for the absence of multicollinearity 
with Variance Inflation Factor (VIF, Table S6, appendix B). The slopes reported
in the main text were standardized in order to compare their magnitude ($r_\delta
= \beta \frac{\sigma_x}{\sigma_y}$, $\beta$ being the unstandardized slope,
$\sigma_x$ and $\sigma_y$ being respectively the standard deviation of the
predictive variable $x$ and of the response variable $y$). The paths whose
slope had a $P$ value inferior or equal to $0.05$ were reported in the main text, all the
path values are presented in Table S7 and S8.

Indirect effects on community biomass and its temporal stability
were computed by multiplying slopes along the path. Only significant direct
slopes were included in the calculation of indirect effects. Total effects were
computed as the sum of direct and indirect effects.
